---
title: "Project 4"
author: "Jiawen Qi (jiq10)"
date: "April 4, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PART A

Find the Moran’s I and Geary’s C for the areas of polygons in the `AlleghenyCounty_Council.shp` and `AlleghenyCounty_Municipal.shp` files by using Rook’s adjacency and Queen’s adjacency. In other words, the attribute of interest here is the area of polygons. Submit a table summarizing the results and their interpretation.

Results Table: 建立表格并解释

# PART B

Perform the global G statistic on the Crime_PA2002.shp file. Submit a table summarizing the results and their interpretation.

Results Table: 建立表格并解释

Perform the Geographically Weighted Regression (GWR) to calculate the `total number of crimes(Index01)` as a function of `Population(Pop_CRI01)`, `Number of crime agencies(AG_CRI01)`, and `area` for Mifflin County in Crime_PA2002.shp file. Weights are based on the inverse Euclidean distance between polygons’ centroid. Submit the regression equation.

# PART C

Submit:
A table showing the interpolated values and the map of interpolated surface using the IDW technique. [10 points]
A table showing the interpolated values and the map of interpolated surface using the OK technique. [30 points]
What are the differences between the two interpolation techniques and between the two interpolated surface maps. [10 points]

# Code

```{r, results='hide', warning=FALSE, tidy=TRUE}
##### Load Libraries #####
library(rgdal) 
library(UScensus2010) 
library(spdep)
library(GWmodel) 
library(spgwr)
library(dplyr)
library(fields)
library(gstat)
library(automap)


##### PART A #####
# Import files
AlleghenyCountyCouncil <- readOGR(".","AlleghenyCounty_Council") # import AlleghenyCounty_Council shapefile
AlleghenyCountyMunicipal <- readOGR(".","AlleghenyCounty_Municipal") # import AlleghenyCounty_Municipal shapefile

# ACC is the abbr of AlleghenyCountyCouncil; ACM is the abbr of AlleghenyCountyMunicipal
ACC.area = areaPoly(AlleghenyCountyCouncil) # calculate the area of each polygon in ACC
ACM.area = areaPoly(AlleghenyCountyMunicipal) # calculate the area of each polygon in ACM

# # construct neighbours list from polygon
ACC.queen <- poly2nb(AlleghenyCountyCouncil, queen = TRUE) 
ACC.rook <- poly2nb(AlleghenyCountyCouncil, queen = FALSE)
ACM.queen <- poly2nb(AlleghenyCountyMunicipal, queen = TRUE)
ACM.rook <- poly2nb(AlleghenyCountyMunicipal, queen = FALSE)

# change neighbour list to listw (get the spatial weights for neighbours lists)
ACC.nb2listw.queen <- nb2listw(ACC.queen)
ACC.nb2listw.rook <- nb2listw(ACC.rook)
ACM.nb2listw.queen <- nb2listw(ACM.queen)
ACM.nb2listw.rook <- nb2listw(ACM.rook)

# calculate moran's and geary'c
moran.test(ACC.area, ACC.nb2listw.queen, randomisation = FALSE) # ACC moran queen: 0.31108273, Expectation = -0.08333333
moran.test(ACC.area, ACC.nb2listw.rook, randomisation = FALSE) # ACC moran rook: 0.31108273, Expectation = -0.08333333
geary.test(ACC.area, ACC.nb2listw.queen, randomisation = FALSE) # ACC geary queen: 0.6409275, Expectation = 1;
geary.test(ACC.area, ACC.nb2listw.rook, randomisation = FALSE) # ACC geary rook: 0.6409275, Expectation = 1;
moran.test(ACM.area, ACM.nb2listw.queen, randomisation = FALSE) # ACM moran queen: -0.0001232697, Expectation = -0.007751938;
moran.test(ACM.area, ACM.nb2listw.rook, randomisation = FALSE) # ACM moran rook: -0.002915556, Expectation = -0.007751938
geary.test(ACM.area, ACM.nb2listw.queen, randomisation = FALSE) # ACM geary queen: 2.09170840, Expectation = 1;
geary.test(ACM.area, ACM.nb2listw.rook, randomisation = FALSE) # ACM geary rook: 2.134989922, Expectation = 1;

##### PART B #####
# Import dataset Crime_PA2002 shapefile
Crime_PA2002 <- readOGR(".","Crime_PA2002")

# Construct neighbours list from polygon list
Crime_PA2002.rook <- poly2nb(Crime_PA2002, queen = FALSE) # rook

# Spatial Weights for neighbours lists
Crime_PA2002.rook.nb2listw <- nb2listw(Crime_PA2002.rook)

##### Global G statistics
globalG.test(Crime_PA2002$BURG01, Crime_PA2002.rook.nb2listw) # Global G statistic: 2.974714e-02 ; Expectation: 1.515152e-02; Variance: 1.600546e-05

##### GWR
# Get the Euclidian Distance weight Matrix for all combinations
DistanceMatrix <- gw.dist(dp.locat=coordinates(Crime_PA2002)) # the euclidian distance matrix
DistanceMatrixMifflin <- DistanceMatrix[61,]
DistanceMatrixMifflin <- DistanceMatrixMifflin[-61]
weight <- 1/DistanceMatrixMifflin
weight

# spgwr package gwr
bw <-gwr.sel(INDEX01 ~ POP_CRI01 + AG_CRI01 + Area, data = Crime_PA2002[Crime_PA2002$COUNTY != 'Mifflin County',])
gwr.model <- gwr(INDEX01~POP_CRI01+AG_CRI01+Area, data = Crime_PA2002[Crime_PA2002$COUNTY != 'Mifflin County',], bandwidth=bw, hatmatrix = TRUE, se.fit = TRUE, weights = weight)
gwr.model
x <- gwr(INDEX01~POP_CRI01+AG_CRI01+Area, data = Crime_PA2002, bandwidth = bw, predict=TRUE, se.fit=TRUE, fittedGWRobject=gwr.model)
x$SDF$pred[61] # the predict result

##### PART C #####
PA_County_Select <- readOGR(".","PA_County_Select")
Ozone_Value <- read.delim("Ozone_Value.dat", header = FALSE, sep = "|")
Ozone_Sensor_Locs <- readOGR(".", "Ozone_Sensor_Locs")
PA_County_Select@data$COUNTY<-as.character(PA_County_Select@data$COUNTY)
names(Ozone_Value)[3]<-"id"

## Inverse Distance Weighting (IDW)
centroids <- as.data.frame(coordinates(PA_County_Select)) # get the centroids for each county
names(centroids) <- c("centroidX", "centroidY") # rename the column name
rownames(centroids) <- seq(1:length(centroids$centroidX)) # rename the row index
ControlPoints <- cbind.data.frame(Ozone_Sensor_Locs$id,Ozone_Sensor_Locs$long, Ozone_Sensor_Locs$lat) # get all the points long and lat that would be the control points
names(ControlPoints) <- c("id","controlX", "controlY") # rename the column names
Sub_Ozone_Value <- Ozone_Value[Ozone_Value$id %in% ControlPoints$id, c(3, 6, 8)] # get the ozone value with the same id in the control points
# Create the IDW calculation table in loop and record the results
Sub_Ozone_Value <- Sub_Ozone_Value[Sub_Ozone_Value$V6 == "OZONE",-2] # get the ozone value
ControlPoints$id <- as.character(ControlPoints$id) # change id to char
Sub_Ozone_Value$id <- as.character(Sub_Ozone_Value$id) # change id to char
ControlPoints <- inner_join(ControlPoints, Sub_Ozone_Value) # Join to get the ozone value for each sensor station
names(ControlPoints)[4] <- "ozone_value"
ResultTable <- as.data.frame(PA_County_Select@data$COUNTY) # This Table will record the results
ResultTable$x <- centroids$centroidX # long of centroid for each polygon
ResultTable$y <- centroids$centroidY # lat of centroid for each polygon
ResultTable$z <- NA # interpolated ozone value
# using for loop to interpolate for each polygon
for (i in 1:length(centroids$centroidX)) {
  centroidX <- ResultTable$x[i]
  centroidY <- ResultTable$y[i]
  IDWtable <- ControlPoints
  IDWtable$Distance <- sqrt((IDWtable$controlX-centroidX)^2+(IDWtable$controlY-centroidY)^2)
  IDWtable <- IDWtable[order(IDWtable$Distance),]
  rownames(IDWtable) <- seq(1:length(IDWtable$id))
  IDWtable <- IDWtable[1:5,] # keep only 5 nearest
  IDWtable$InverseDistance <- 1/IDWtable$Distance
  IDWtable$Weight <- IDWtable$InverseDistance/sum(IDWtable$InverseDistance)
  IDWtable$WeightedValue <- IDWtable$Weight*IDWtable$ozone_value
  ResultTable$z[i] <- sum(IDWtable$WeightedValue)
}
ResultTable <- ResultTable[,-1]
ResultTable

# Create the IDW Map
IDW.map<-as.data.frame(ResultTable)
coordinates(IDW.map)=~x+y
xRange<-as.numeric(bbox(IDW.map)[1,])
yRange<-as.numeric(bbox(IDW.map)[2,])
grid<-expand.grid(x = seq(from = xRange[1], to =xRange[2], by = 0.01),
                  y = seq(from = yRange[1],to = yRange[2], by = 0.01))
coordinates(grid) <- ~x + y
gridded(grid) <- TRUE
IDW.data <- gstat::idw(ResultTable$z ~ 1, locations=IDW.map, newdata=grid)
OzonePlot<- par(mar=c(0,0,0,0))
image(IDW.data,"var1.pred",col=terrain.colors(50))
contour(IDW.data,"var1.pred", add=TRUE, nlevels=10)
plot(IDW.map, add=TRUE, pch=10, cex=0.5)
text(coordinates(PA_County_Select),
     paste(as.character(round(ResultTable$z,1)),
           as.character(PA_County_Select$COUNTY) ), 
     pos=4, cex=0.8, col="black")
map.axes(cex.axis=0.8)
par(OzonePlot)

## Ordinary Kriging (OK) 
Ozone_Value <- Ozone_Value[Ozone_Value$V6 == 'OZONE', ] # get all the ozone value
Ozone_Sensor_Locs_data <- Ozone_Sensor_Locs@data
Ozone_Sensor_Locs_data$id <- as.character(Ozone_Sensor_Locs_data$id)
Ozone_Value$id <- as.character(Ozone_Value$id)
Ozone_Sensor_Value <- inner_join(x = Ozone_Sensor_Locs_data, y = Ozone_Value) # join by id
Sensor.num <- nrow(Ozone_Sensor_Value) # there are 11 sensors has value
Ozone_Sensor_Value <- subset.data.frame(Ozone_Sensor_Value, select = c(long, lat, V8))
names(Ozone_Sensor_Value) <- c('x', 'y', 'z')
centroids$z <- NA # create the new column z to store the ozone value
names(centroids) <- c('x', 'y', 'z')
dataset <- rbind.data.frame(Ozone_Sensor_Value, centroids)

# Create matrix D of distance between control points
D <- as.data.frame(matrix(data = NA, nrow = Sensor.num, ncol = Sensor.num))
names(D) <- c(1:nrow(D))
for (i in 1:nrow(D)) {
  for (j in 1:nrow(D)) {
    D[i,j] = sqrt((Ozone_Sensor_Value$x[i] - Ozone_Sensor_Value$x[j])^2 + (Ozone_Sensor_Value$y[i] - Ozone_Sensor_Value$y[j])^2)
  }
}

# exponential semivariogram model
variogram = Ozone_Sensor_Value 
coordinates(variogram) = ~x + y 
variogram <- autofitVariogram(z~x+y, variogram, model="Exp") # exponential
plot(variogram) # sill is 15, range is 0.11, nugget is 14
a <- 0.11 # range
c <- 15 # sill
n <- 14
gamma <- function(d) {
  g <- n + (c - n) * (1-exp(-d/a))
  return(g)
}

# get Matrix A
A <- gamma(D)
A <- rbind.data.frame(A, 1)
A <- cbind.data.frame(A, 1)
diag(A) <- 0
names(A) <- c(seq(1:12))
A.inverse <- solve(A)

for (i in 1:10) { # for each centroid
  # get vector d, d
  d <- c()
  for (j in 1:11) { # for each control point
    distance <- sqrt((centroids$x[i] - Ozone_Sensor_Value$x[j]) ^ 2 + (centroids$y[i] - Ozone_Sensor_Value$y[j]) ^ 2)
    d <- c(d, distance)
  }
  # get vector b
  b <- c(gamma(d),1)
  # get the weight w
  w = A.inverse %*% b
  z.estimate <- sum(ControlPoints$ozone_value * w[1:11])
  centroids$z[i] = z.estimate
}
centroids

# create the OK map  // parameters
OK.result <- as.data.frame(centroids)
coordinates(OK.result) = ~ x + y
variogram <- autofitVariogram(z ~ x + y, OK.result, model = "Exp")
variogram
plot(variogram)
nugget = 0.01
sill = 0.02
range = 0.17
model <- vgm(psill = sill, model = "Exp", range = range, nugget = nugget)
krige <- krige(OK.result$z ~ 1, OK.result, grid, model = model) # using ordinary kriging
OzonePlot2 <- par(mar = c(0, 0, 0, 0))
image(krige ,"var1.pred",col=terrain.colors(50)) 
contour(krige ,"var1.pred", add=TRUE, nlevels=10) 
plot(OK.result, add=TRUE, pch=10, cex=0.5)
text(coordinates(OK.result), 
     paste(as.character(round(OK.result$z, 1)),
           as.character(PA_County_Select$COUNTY)), 
     pos=4, 
     cex=0.8,
     col="black")
map.axes(cex.axis=0.8)
par(OzonePlot2)



```